### 2018-08-08 ZCB Pricing .Rmd
Comment changes

---
title: "ZCB Pricing"
author:
- Valentine Chisango
- George Parekkadavil
- Vegan Pather
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls(all =TRUE))
```

```{r 1}
# Closed form solution of the Bond price

bond_price <- function(alpha=0.08,beta=0.8,sigma=0.005,t=0,T=5,r=0.08){
  # alpha, beta, sigma <- Fixed Parameters
  # T <- maturity time
  # r <- Initial rate at time 0  
  A <- (1-exp(-beta*(T-t)))/(beta)
  B <- (1/beta)*((alpha-(sigma)^2/(2*beta))*(A - T +t) - (sigma^(2)*A^2)/(4))
  exp(-A*r +B)
}

```

```{r 2}

# Crude Monte Carlo function

Crude_Monte_Carlo=function(alpha =0.08,beta=0.8,sigma=0.005,T=5,r = 0.08,n=1000,s=100){

  # alpha, beta, sigma <- Fixed Parameters
  # T <- maturity time
  # r <- Initial rate at time 0
  # n <- Number of Simulations
  # s <- NUmber of Sub-intervals over time

  #Simulate interest rates
  dt <- T/s # Length of sub-intervals
  sim_path = matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path[1,]= r # Initial rates on first row.

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*rnorm(1)# Discretize the SDE of the interest model using Euler Discretization to simulate rates over discrete sub-intervals
      sim_path[j,i]=sim_path[j-1,i]+dr
    }
  }

  # Monte Carlo Simulations
  Sum = colSums(sim_path[2:(s+1),]*dt) # Approximate the integral using summation
  int = exp(-Sum) # Take the exponent of the Integral
  mean = mean(int) # Find the average of expression
  mean

}

```

```{r 3}
# Improvements to the Monte Carlo Simulation

# Antithetic Variates

Antithetic_variates = function(alpha =0.08,beta=0.8,sigma=0.005,T=5,r = 0.08,n=1000,s=100){

  # alpha, beta, sigma <- Fixed Parameters
  # T <- maturity time
  # r <- Initial rate at time 0
  # n <- Number of Simulations
  # s <- NUmber of Sub-intervals over time

  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path_inv= matrix(0,nrow = s+1, ncol = n)# matrix to generate interest paths using negative values of RV
  sim_path[1,] = r
  sim_path_inv[1,] = r
  z = rnorm(n) # Generate standard normal values
  z. = -z # Negative RV

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr = (alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*z[i] # Same Euler Discretization as in the Crude Monte Carlo function
      dr_i = (alpha-beta*sim_path_inv[j-1,i])*dt + sigma*sqrt(dt)*z.[i] # Euler Discretization for the use of -Z_i
      sim_path[j,i] =sim_path[j-1,i]+dr
      sim_path_inv[j,i] = sim_path_inv[j-1,i]+dr_i

    }
  }

  # Monte Carlo Simulations
  Sum = colSums(sim_path[2:(s+1),]*dt) # The Sum and Sum. variables record the summation of their respective rates to approximate the integral
  Sum.= colSums(sim_path_inv[2:(s+1),]*dt)
  int = exp(-Sum.) #Exponential of the estimate using Z_i
  int.= exp(-Sum) #Exponential of the estimate using -Z_i
  mean_1 = mean(int) # Estimate using Z_i
  mean_2 = mean(int.)  # Estimate using -Z_i
  result = (mean_1+mean_2)/2 # Final estimate is taken as the average of the two estimates.
  result

}

```

```{r 4}

# Moment matching

Moment_Matching = function(alpha = 0.08, beta = 0.8, sigma = 0.005, T = 5,r = 0.08, n = 1000, s = 200){

  # alpha, beta, sigma <- Fixed Parameters
  # T <- maturity time
  # r0 <- Initial rate at time 0
  # n <- Number of Simulations
  # s <- NUmber of Subintervals

  #Simulate interest rates
  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path[1,]= r
  z=rnorm(n)
  z_m= z-mean(z) # Moment matching transformation of Z_i to match finite moments

  # Follow the same process as in the Monte Carlo approach , but use the transformed Z_i's to simulate the rates
  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*z_m[i] 
      sim_path[j,i]=sim_path[j-1,i]+dr
    }
  }

  Sum=colSums(sim_path[2:(s+1),]*dt)
  int=exp(-Sum)
  result=mean(int)
  result

}

```

```{r 5}

# Control Variates

# Make use of the CIR model as the control variate process.
# Below is the closed form solution under the CIR model
bond_price_CIR <- function(alpha =0.08,beta=0.8,sigma=0.005,t=0,T=5,r = 0.08){
  h <- sqrt(beta^(2)+2*sigma^(2))
  A <-(2*h*exp((beta+h)*(T-t)/2)/(2*h+(beta+h)*(exp((T-t)*(h))-1)))^((2*alpha)/sigma^(2))
  B <-2*(exp((T-t)*(h))-1)/(2*h+(beta+h)*(exp((T-t)*(h))-1))
  A*exp(-B*r)
}

# Monte Carlo Simulations of Bond Price under the CIR model

CIR_MC= function(alpha =0.08,beta=0.8,sigma=0.005,T=5,r = 0.08,n=1000,s=100){

  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path[1,]= r

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+(sigma*sqrt(dt)*rnorm(1)*sqrt(sim_path[j-1,i]))# Euler Discretization
      sim_path[j,i]=sim_path[j-1,i]+dr

    }
  }

  # Monte Carlo Simulations
  Sum=colSums(sim_path[2:(s+1),]*dt)
  int=exp(-Sum)
  result=mean(int)
  result

}
# The control variate estimate is taken as the difference between the the Crude MOnte carlo estimates of the Vasicek and CIR model plus the closed form of the CIR model

Control_Variates_estimate= function(alpha =0.08,beta=0.8,sigma=0.005,t=0,T=5,r = 0.08,n=1000,s=100){
  Crude_Monte_Carlo(alpha,beta,sigma,T,r,n,s)- CIR_MC(alpha,beta,sigma,T,r,n,s)+bond_price_CIR(alpha,beta,sigma,t,T,r)
}

```

```{r 6}
# Explicit Method

explicit = function(alpha = 0.08,beta = 0.8, sigma = 0.005,r = 0.08, I = 1:1250, J = seq(-0.4,0.6,0.005),T = 5){

  # I = partition of time interval
  # J = partition of short rates sequence

  M <- length(J) # Number of short rate grid points
  N <- length(I) # Number of time grid poiints

  dt <- T/N # Lenght of each time subinterval
  dr <- J[2] -J[1] # Lenght of each time interval

  # For each j in J we define the following functions at a fixed time point

  pd <- function(j){
    ((J[j]+1/dt)^(-1))*(sigma^2/(2*(dr^2)) -(alpha -beta*J[j])/(2*dr))
  }

  pm <- function(j){
    ((J[j]+1/dt)^(-1))*(1/dt -sigma^2/(dr^2))
  }

  pu <- function(j){
    ((J[j]+1/dt)^(-1))*(sigma^2/(2*(dr^2)) + (alpha -beta*J[j])/(2*dr))
  }

  #Setting up tridiagonal matrix

  Diag_1 <- matrix(0,ncol = M, nrow = M)

  # Filling main diagonal of matrix 
  for (j in 1:M){
    Diag_1[j,j] <- pm(j)

  }

  # Filling lower diagonal of matrix 
  for (j in 2:M){
    Diag_1[j,j-1] <- pd(j)
  }

  # Filling upper diagonal of matirx 
  for (j in 1:(M-1)){
    Diag_1[j,j+1] <- pu(j)
  }

  # Bond price values at maturity are all 1 unit   
  Bond_values <- matrix(1,nrow=M, ncol =1)

  # Working backwards in time
  for (i in 1:(length(I))){
    C <- matrix(0,ncol = 1, nrow = M) # Another matrix for boundary conditions
    C[M,1] <- (pu(M))*Bond_values[M,1]
    C[1,1] <- pd(1)*Bond_values[1,1]

    Bond_values <- Diag_1%*%Bond_values
    Bond_values <- Bond_values
  }

  # Finding index of the desired short rate
  index = match(as.character(r),J)
  # Output is the corresponding bond price at time 0
  Bond_values[index]
}

```

```{r 7}
# Implicit Method

implicit <- function(alpha = 0.08,beta = 0.8,sigma = 0.005,r = 0.08,I = 1:1250, J =seq(-0.4,0.6,0.005),T=5){

  # I = partition of time interval
  # J = partition of short rate variable

  M <- length(J) # Number of short rate grid points
  N <- length(I) # Number of time grid poiint

  dr <- J[2] - J[1] # length of each sub-interval across J
  dt <- T/N # length of each sub-interval across I

  # For each j in J we define the following functions at a fixed time point

  pd <- function(j){
    dt*((alpha-beta*J[j])/(2*dr) - (sigma^2)/(2*dr^2))

  }

  pm <- function(j){
    dt*(J[j] + 1/dt +sigma^2/(dr^2))
  }

  pu <- function(j){
    -dt*((alpha-beta*J[j])/(2*dr) + (sigma^2)/(2*dr^2))
  }

  # Setting up tridiagonal matrix
  Diag <- matrix(0,nrow = M,ncol = M)

  # Filling main diagonal
  for (j in 1:M){
    Diag[j,j] <- pm(j)

  }

  # Filling lower diagonal
  for (j in 2:M){
    Diag[j,j-1] <- pd(j)
  }

  # Filling upper diagonal
  for (j in 1:(M-1)){
    Diag[j,j+1] <- pu(j)
  }

  Bond_values <- matrix(1,ncol = 1,nrow = M) # The matrix of bond prices at maturity time 
  Inv <- solve(Diag) # Inverse of tridiagonal matrix

  for (i in 1:(length(I))){
    C <- matrix(0,ncol = 1, nrow = M) # Another matrix for boundary conditions
    C[M,1] <- (-pu(M))*Bond_values[M,1]
    C[1,1] <- -pd(1)*Bond_values[1,1]

    RHS <- Bond_values + C # Working backwards recursively 
    Bond_values <- Inv%*%RHS
  }

  # Finding index of the desired short rate
  index = match(as.character(r),J)
  # Output is the corresponding bond price at time 0
  Bond_values[index]
}

```

```{r 8}
# Crank-Nicholson Method

CN <- function(alpha = 0.08, beta = 0.8, sigma = 0.005, r = 0.08, T = 5 , I = 1:1250, J = seq(-0.4,0.6,0.005)){

  # I = partition of time interval
  # J = partition of short rate interval

  M <- length(J) # Number of short rate grid points
  N <- length(I) # Number of time grid poiints

  dt <- T/N # Length of each time subinterval
  dr <- J[2] -J[1] # Lenght of each rate subinterval

  #========================================================================================
  # Define the following functions for j = 1,...,M at each time point i

  pd <- function(j){
    (alpha -beta*J[j])/(4*dr) - sigma^2/(4*(dr^2))
  }

  pm <- function(j){
    sigma^2/(2*(dr^2)) + J[j]/2
  }

  pu <- function(j){
    -(alpha -beta*J[j])/(4*dr) - sigma^2/(4*(dr^2)) 
  }

  #========================================================================================
  # Define 2 tridiagonal matricies. One for time point i and one for time point i+1

  Diag_1 <- matrix(0,nrow = M,ncol = M)
  Diag_2 <- matrix(0,nrow = M,ncol = M)

  # Filling main diagonal of matrix 1
  for (j in 1:M){
    Diag_1[j,j] <- -1/dt - pm(j)
  }

  # Filling lower diagonal of matrix 1
  for (j in 2:M){
    Diag_1[j,j-1] <- -pd(j)
  }

  # Filling upper diagonal of matirx 1
  for (j in 1:M-1){
    Diag_1[j,j+1] <- -pu(j)
  }

  # Filling main diagonal of matrix 2
  for (j in 1:M){
    Diag_2[j,j] <- -1/dt + pm(j)
    }

  # Filling lower diagonal of matrix 2
  for (j in 2:M){
    Diag_2[j,j-1] <- pd(j)
    }

  # Filling upper diagonal of matirx 2
  for (j in 1:M-1){
    Diag_2[j,j+1] <- pu(j)
  }

  #========================================================================================
  # Constructing more matricies and iterating the procedure from maturity to time 0
  Inv <- solve(Diag_1)

  Bond_values_1 <- matrix(1,nrow = M, ncol = 1)
  Bond_values_2 <- matrix(1,nrow = M, ncol = 1)
  C_1 <- matrix(0,nrow = M,ncol = 1)
  C_1[1,1] <- pd(1)*Bond_values_1[1,1]
  C_1[M,1] <- pu(M)*Bond_values_1[M,1]
  RHS <- Diag_2%*%Bond_values_1 +2*C_1
  Bond_values_1 <- Inv%*%RHS

  for (i in 1:(length(I)-1)){
    C_2 <- matrix(0,nrow = M,ncol = 1)
    C_2[1,1] <- pd(1)*Bond_values_2[1,1]
    C_2[M,1] <- pu(M)*Bond_values_2[M,1]

    C_1 <- matrix(0,nrow = M,ncol = 1)
    C_1[1,1] <- pd(1)*Bond_values_1[1,1]
    C_1[M,1] <- pu(M)*Bond_values_1[M,1]

    RHS <- Diag_2%*%Bond_values_1 + C_1 + C_2
    Bond_values <- Inv%*%RHS
    Bond_values_2 <- Bond_values_1
    Bond_values_1 <- Bond_values
  }

  # Index of rate corresponding to desired short rate
  index = match(as.character(r), J)
  # Output corresponding to desired short rate
  Bond_values[index]
}

```

```{r 9}
# Plots of the different methods
par(mfrow = c(2,2))

xvec <- seq(-0.4,0.6,0.005) # Vector for rates

known_sol <- NULL # Vector to hold the known closed form solution
for (r in 1:length(xvec)){
  known_sol[r] <- bond_price(0.08,0.8,0.005,0,5,xvec[r])
}

CR <- NULL # Vector to hold Crude Monte Carlo Solutions
AV <- NULL # Vector to hold Antithetic Monte Carlo Solutions
MM <- NULL # Vector to hold Moment Matching  Carlo Solutions
CV <- NULL # Vector to hold Control Variates Monte Carlo Solutions

Expl <- NULL # Vector to hold Explicit method Solutions
Impl <- NULL # Vector to hold Implicit method Solutions
Crank <- NULL # Vector to hold Crank-Nicolson Solutions

# Plots of Monte Carlo Methods ----------------------------------------------------------------------

for (i in 1: length(xvec)){
  CR[i] <- Crude_Monte_Carlo(r = xvec[i])
}

for (i in 1: length(xvec)){
  AV[i] <- Antithetic_variates(r = xvec[i])
}

for (i in 1: length(xvec)){
  MM[i] <- Moment_Matching(r = xvec[i])
}

for (i in 1: length(xvec)){
  CV[i] <- Control_Variates_estimate(r = xvec[i])
}

plot(CR, x = xvec[1:length(xvec)],xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Monte Carlo Solutions vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)
lines(xvec, AV, col ="black", lty =1.5)
lines(xvec, MM, col ="yellow", lty =1.5)
lines(xvec, CV, col ="green", lty =1.5)

# Plots of Finite Difference Methods ----------------------------------------------------------------------

for (i in 1: length(xvec)){
  Expl[i] <- explicit(r = xvec[i], I = 1:10000)
}

for (i in 1: length(xvec)){
  Impl[i] <- implicit(r = xvec[i])
}

for (i in 1: length(xvec)){
  Crank[i] <- CN(r = xvec[i])
}

plot(Expl, x = xvec ,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Explicit Method vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)

plot(Impl, x = xvec ,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Implicit Method vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)

plot(Crank, x = xvec ,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Crank-Nicolson Method vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)

```

```{r 10}
# Comparison of results for numerical methods

compare_mat <- matrix(NA,nrow = 7, ncol = 4) # matrix of comparison

colnames(compare_mat) <- c("Price", "Known Price", "Error", "System Time")
rownames(compare_mat) <- c("Crude Monte Carlo", "Antithetic Variates", "Moment Matching", "Control Variates", "Explicit Method", "Implicit Method", "Crank-Nicolson Method")

compare_mat[,2] <- bond_price()

# Crude Monte Carlo
start_time <- proc.time()[1]
compare_mat[1,1] <- Crude_Monte_Carlo()
compare_mat[1,4] <- proc.time()[1] - start_time
compare_mat[1,3] <- compare_mat[1,2] - compare_mat[1,1]

# Antithetic Variates
start_time <- proc.time()[1]
compare_mat[2,1] <- Antithetic_variates()
compare_mat[2,4] <- proc.time()[1] - start_time
compare_mat[2,3] <- compare_mat[2,2] - compare_mat[2,1]

# Moment Matching
start_time <- proc.time()[1]
compare_mat[3,1] <- Moment_Matching()
compare_mat[3,4] <- proc.time()[1] - start_time
compare_mat[3,3] <- compare_mat[3,2] - compare_mat[3,1]

# Control Variates
start_time <- proc.time()[1]
compare_mat[4,1] <- Control_Variates_estimate()
compare_mat[4,4] <- proc.time()[1] - start_time
compare_mat[4,3] <- compare_mat[4,2] - compare_mat[4,1]

# Explicit method
start_time <- proc.time()[1]
compare_mat[5,1] <- explicit()
compare_mat[5,4] <- proc.time()[1] - start_time
compare_mat[5,3] <- compare_mat[5,2] - compare_mat[5,1]

# Implicit method
start_time <- proc.time()[1]
compare_mat[6,1] <- implicit()
compare_mat[6,4] <- proc.time()[1] - start_time
compare_mat[6,3] <- compare_mat[6,2] - compare_mat[6,1]

# Crank-Nicolson method
start_time <- proc.time()[1]
compare_mat[7,1] <- CN()
compare_mat[7,4] <- proc.time()[1] - start_time
compare_mat[7,3] <- compare_mat[7,2] - compare_mat[7,1]

library(xtable)
xtable(compare_mat, digits = 5)

```

```{r 11}
# Sensitivity Tests

senses <- matrix(nrow=13, ncol=9) # matrix to hold prices for various parameters
colnames(senses) <-c("Parameters","Closed Form","Crude Monte Carlo","Antithetic Variates","Moment Matching","Control Variates","Explicit","Implicit","Crank-Nicholson")
xvec=c("Closed Form","Crude","AV","MM","CV","Explicit","Implicit","CN") #list of numerical methods

#Initial Parameters
senses[1,2]=bond_price()
senses[1,3]=Crude_Monte_Carlo()
senses[1,4]=Antithetic_variates()
senses[1,5]=Moment_Matching()
senses[1,6]=Control_Variates_estimate()
senses[1,7]=explicit()
senses[1,8]=implicit()
senses[1,9]=CN()

#r=0.05
senses[2,2]=bond_price(r=0.05)
senses[2,3]=Crude_Monte_Carlo(r=0.05)
senses[2,4]=Antithetic_variates(r=0.05)
senses[2,5]=Moment_Matching(r=0.05)
senses[2,6]=Control_Variates_estimate(r=0.05)
senses[2,7]=explicit(r=0.05)
senses[2,8]=implicit(r=0.05)
senses[2,9]=CN(r=0.05)

#r=0.1
senses[3,2]=bond_price(r=0.1)
senses[3,3]=Crude_Monte_Carlo(r=0.1)
senses[3,4]=Antithetic_variates(r=0.1)
senses[3,5]=Moment_Matching(r=0.1)
senses[3,6]=Control_Variates_estimate(r=0.1)
senses[3,7]=explicit(r=0.1)
senses[3,8]=implicit(r=0.1)
senses[3,9]=CN(r=0.1)

#T=1
senses[4,2]=bond_price(T=1)
senses[4,3]=Crude_Monte_Carlo(T=1)
senses[4,4]=Antithetic_variates(T=1)
senses[4,5]=Moment_Matching(T=1)
senses[4,6]=Control_Variates_estimate(T=1)
senses[4,7]=explicit(T=1)
senses[4,8]=implicit(T=1)
senses[4,9]=CN(T=1)

#T=10
senses[5,2]=bond_price(T=10)
senses[5,3]=Crude_Monte_Carlo(T=10)
senses[5,4]=Antithetic_variates(T=10)
senses[5,5]=Moment_Matching(T=10)
senses[5,6]=Control_Variates_estimate(T=10)
senses[5,7]=explicit(T=10)
senses[5,8]=implicit(T=10)
senses[5,9]=CN(T=10)

#sigma=0.001
senses[6,2]=bond_price(sigma=0.001)
senses[6,3]=Crude_Monte_Carlo(sigma=0.001)
senses[6,4]=Antithetic_variates(sigma=0.001)
senses[6,5]=Moment_Matching(sigma=0.001)
senses[6,6]=Control_Variates_estimate(sigma=0.001)
senses[6,7]=explicit(sigma=0.001)
senses[6,8]=implicit(sigma=0.001)
senses[6,9]=CN(sigma=0.001)

#sigma=0.01
senses[7,2]=bond_price(sigma=0.01)
senses[7,3]=Crude_Monte_Carlo(sigma=0.01)
senses[7,4]=Antithetic_variates(sigma=0.01)
senses[7,5]=Moment_Matching(sigma=0.01)
senses[7,6]=Control_Variates_estimate(sigma=0.01)
senses[7,7]=explicit(sigma=0.01)
senses[7,8]=implicit(sigma=0.01)
senses[7,9]=CN(sigma=0.01)

#alpha=0.01
senses[8,2]=bond_price(alpha=0.01)
senses[8,3]=Crude_Monte_Carlo(alpha=0.01)
senses[8,4]=Antithetic_variates(alpha=0.01)
senses[8,5]=Moment_Matching(alpha=0.01)
senses[8,6]=Control_Variates_estimate(alpha=0.01)
senses[8,7]=explicit(alpha=0.01)
senses[8,8]=implicit(alpha=0.01)
senses[8,9]=CN(alpha=0.01)

#alpha=0.16
senses[9,2]=bond_price(alpha=0.16)
senses[9,3]=Crude_Monte_Carlo(alpha=0.16)
senses[9,4]=Antithetic_variates(alpha=0.16)
senses[9,5]=Moment_Matching(alpha=0.16)
senses[9,6]=Control_Variates_estimate(alpha=0.16)
senses[9,7]=explicit(alpha=0.16)
senses[9,8]=implicit(alpha=0.16)
senses[9,9]=CN(alpha=0.16)

#beta=0.1
senses[10,2]=bond_price(beta=0.1)
senses[10,3]=Crude_Monte_Carlo(beta=0.1)
senses[10,4]=Antithetic_variates(beta=0.1)
senses[10,5]=Moment_Matching(beta=0.1)
senses[10,6]=Control_Variates_estimate(beta=0.1)
senses[10,7]=explicit(beta=0.1)
senses[10,8]=implicit(beta=0.1)
senses[10,9]=CN(beta=0.1)

#beta=1.6
senses[11,2]=bond_price(beta=1.6)
senses[11,3]=Crude_Monte_Carlo(beta=1.6)
senses[11,4]=Antithetic_variates(beta=1.6)
senses[11,5]=Moment_Matching(beta=1.6)
senses[11,6]=Control_Variates_estimate(beta=1.6)
senses[11,7]=explicit(beta=1.6)
senses[11,8]=implicit(beta=1.6)
senses[11,9]=CN(beta=1.6)

#alpha=0.01,beta=0.1
senses[12,2]=bond_price(alpha=0.01,beta=0.1)
senses[12,3]=Crude_Monte_Carlo(alpha=0.01,beta=0.1)
senses[12,4]=Antithetic_variates(alpha=0.01,beta=0.1)
senses[12,5]=Moment_Matching(alpha=0.01,beta=0.1)
senses[12,6]=Control_Variates_estimate(alpha=0.01,beta=0.1)
senses[12,7]=explicit(alpha=0.01,beta=0.1)
senses[12,8]=implicit(alpha=0.01,beta=0.1)
senses[12,9]=CN(alpha=0.01,beta=0.1)

#alpha=0.16,beta=1.6
senses[13,2]=bond_price(alpha=0.16,beta=1.6)
senses[13,3]=Crude_Monte_Carlo(alpha=0.16,beta=1.6)
senses[13,4]=Antithetic_variates(alpha=0.16,beta=1.6)
senses[13,5]=Moment_Matching(alpha=0.16,beta=1.6)
senses[13,6]=Control_Variates_estimate(alpha=0.16,beta=1.6)
senses[13,7]=explicit(alpha=0.16,beta=1.6)
senses[13,8]=implicit(alpha=0.16,beta=1.6)
senses[13,9]=CN(alpha=0.16,beta=1.6)

par(mfrow=c(1,1))
plot(x=1:8,y=senses[1,2:9], ylim =c(0.58,0.66), xlab = "Numerical Method", ylab = "Bond Price", main = "Sensitivity Test: Interest Rate", xaxt = "n", col ="black", pch = 19, font.main = 1)
axis(1, at=1:8, labels=xvec)
lines(1:8, senses[2,2:9], col="red",type = "p", pch = 19)
lines(1:8, senses[3,2:9],col ="blue",type = "p", pch =19)
legend("bottomright", inset = c(-0,0), legend = c("r = 0.05", "r = 0.08", "r = 0.10"), lty = c(1,1,1), col = c("red","black","blue", "red"), cex = 0.8, bty ="n", pch =19)

plot(x=1:8,y=senses[1,2:9], ylim =c(0,2.2), xlab = "Numerical Method", ylab = "Bond Price", main = "Sensitivity Test: Time to maturity")
#axis(1, at=1:8, labels=xvec)
lines(1:8, senses[4,2:9], col="red",type = "p")
lines(1:8, senses[5,2:9],col ="blue",type = "p")
legend("topright", inset = c(-0,0), legend = c("T=1", "T=5", "T=10"), lty = c(1,1,1), col = c("red","black","blue", "red"), cex = 0.8, bty ="n")

plot(x=1:8,y=senses[1,2:9], ylim =c(0,2.2), xlab = "Numerical Method", ylab = "Bond Price", main = "Sensitivity Test: Sigma")
#axis(1, at=1:8, labels=xvec)
lines(1:8, senses[6,2:9], col="red",type = "p")
lines(1:8, senses[7,2:9],col ="blue",type = "p")
legend("bottomright", inset = c(-0,0), legend = c("sigma=0.001", "sigma=0.005", "sigma=0.1"), lty = c(1,1,1), col = c("red","black","blue", "red"), cex = 0.8, bty ="n")

plot(x=1:8,y=senses[1,2:9], ylim =c(0,2.2), xlab = "Numerical Method", ylab = "Bond Price", main = "Sensitivity Test: Alpha")
#axis(1, at=1:8, labels=xvec)
lines(1:8, senses[8,2:9], col="red",type = "p")
lines(1:8, senses[9,2:9],col ="blue",type = "p")
legend("bottomright", inset = c(-0,0), legend = c("alpha=0.01", "alpha=0.08", "alpha=0.16"), lty = c(1,1,1), col = c("red","black","blue", "red"), cex = 0.8, bty ="n")

plot(x=1:8,y=senses[1,2:9], ylim =c(0,2.2), xlab = "Numerical Method", ylab = "Bond Price", main = "Sensitivity Test: Beta")
#axis(1, at=1:8, labels=xvec)
lines(1:8, senses[10,2:9], col="red",type = "p")
lines(1:8, senses[11,2:9],col ="blue",type = "p")
legend("topright", inset = c(-0,0), legend = c("beta=0.1", "beta=0.8", "beta=1.6"), lty = c(1,1,1), col = c("red","black","blue", "red"), cex = 0.8, bty ="n")

plot(x=1:8,y=senses[1,2:9], ylim =c(0,2.2), xlab = "Numerical Method", ylab = "Bond Price", main = "Sensitivity Test: Alpha & Beta")
#axis(1, at=1:8, labels=xvec)
lines(1:8, senses[12,2:9], col="red",type = "p")
lines(1:8, senses[13,2:9],col ="blue",type = "p")
legend("topright", inset = c(-0,0), legend = c("alpha=0.01,beta=0.1", "alpha=0.08,beta=0.8", "alpha=0.16,beta=1.6"), lty = c(1,1,1), col = c("red","black","blue", "red"), cex = 0.8, bty ="n")
```

```{r 12}
#Stability Tests

# We do this by checking the variance amongst the results when simulated many times.

CR = NULL
AN = NULL
MM = NULL
CV = NULL

for ( i in 1:100){
  CR[i]=Crude_Monte_Carlo()
  AN[i]=Antithetic_variates()
  MM[i]=Moment_Matching()
  CV[i]=Control_Variates_estimate()
}

```

```{r 13}
# Convergence Tests for finite difference methods
# Varying time steps

time_vec <- seq(1,100,1)

closed <- NULL
CN_conv <- NULL
Impl_conv <- NULL
Expl_conv <- NULL

for (i in 1:length(time_vec)){
  closed[i] <- bond_price()
}

for (i in 1:length(time_vec)){
  CN_conv[i] <- CN(I = 1:time_vec[i])
}

for (i in 1:length(time_vec)){
  Impl_conv[i] <- implicit(I = 1:time_vec[i])
}

for (i in 1:length(time_vec)){
  Expl_conv[i] <- explicit(I = 1:time_vec[i])
}

par(mfrow = c(1,1))
plot(closed, x = time_vec, type = "l", ylim = c(0.59,0.65), xlab = "Number of time steps", ylab = "Estimated Price" , lty =1, font.main = 1, main = "Finite Difference Bond Prices as Time Steps Increase")
lines(x = time_vec, y = CN_conv, type = "l", lty = 2, lwd =1.5, col = "green")
lines(x = time_vec, y = Impl_conv, type = "l", lty = 2, col = "blue",lwd =1.5)
lines(x = time_vec, y = Expl_conv, type = "l", lty = 2, col = "red",lwd =1.5)
legend(x = 60, y = 0.62, legend = c("Closed Form", "Crank-Nicolson", "Implicit", "Explicit"), lty = c(1,2,2,2), col = c("black","green","blue", "red"), cex = 0.8, bty ="n")

```

```{r 14}
# Convergence Tests for finite difference methods
# Varying rate steps

# rate_step <- c(0.02,0.016,0.012,0.010,0.008,0.006,0.005,0.004,0.003,0.002)
# rate_vec <- NULL
# for (i in 1:length(rate_step)){
#   rate_vec[i] <- length(seq(-0.4,0.6,rate_step[i])) 
# }
# 
# 
# closed_r <- NULL
# CN_conv_1 <- NULL
# Impl_conv_1 <- NULL
# Expl_conv_1 <- NULL
# 
# 
# 
# for (i in 1:length(rate_step)){
#   closed_r[i] <- bond_price()
# }
# 
# for (i in 1:length(rate_step)){
#   CN_conv_1[i] <- CN(J = seq(-0.4,0.6,rate_step[i]))
# }
# 
# for (i in 1:length(rate_step)){
#   Impl_conv_1[i] <- implicit(J = seq(-0.4,0.6,rate_step[i]))
# }
# 
# for (i in 1:length(rate_step)){
#   Expl_conv_1[i] <- explicit(J = seq(-0.4,0.6,rate_step[i]))
# }
# 
# par(mfrow = c(1,1))
# plot(closed_r, x = rate_vec, type = "l", xlab = "Number of time steps", ylab = "Estimated Price")
# lines(x = rate_vec, y = CN_conv_1, type = "l", lty = 2)
# lines(x = rate_vec, y = Impl_conv_1, type = "l", lty = 3, col = "blue")
# lines(x = rate_vec, y = Expl_conv_1, type = "l", lty = 4, col = "red")
# legend(x = 60, y = 0.6, legend = c("Closed Form", "Crank-Nicolson", "Implicit", "Explicit"), lty = 1:4, col = c("black","black","blue", "red"), cex = 0.8)

```

```{r 15}
# Covergence tests for Monte Carlo functions

trials_vec <- seq(10,2000,50)

closed_1 <- NULL
CR_conv <- NULL
AV_conv <- NULL
MM_conv <- NULL
CV_conv <- NULL

for (i in 1:length(trials_vec)){
  closed_1[i] <- bond_price()
}

for (i in 1:length(trials_vec)){
  CR_conv[i] <- Crude_Monte_Carlo(n = trials_vec[i])
}

for (i in 1:length(trials_vec)){
  AV_conv[i] <- Antithetic_variates(n = trials_vec[i])
}

for (i in 1:length(trials_vec)){
  MM_conv[i] <- Moment_Matching(n = trials_vec[i])
}

for (i in 1:length(trials_vec)){
  CV_conv[i] <- Control_Variates_estimate(n = trials_vec[i])
}

plot(closed_1, x = trials_vec, type = "l",ylim = c(0.61,0.63), xlab = "Number of Trials", ylab = "Estimated Price", lty = 2, lwd = 2,font.main = 1, main = "Monte Carlo Bond Prices over Number of Trials")
lines(x = trials_vec, y = CR_conv, type = "l", lty = 1, lwd =1.5)
lines(x = trials_vec, y = AV_conv, type = "l", lty = 1, col = "blue", lwd =1.5)
lines(x = trials_vec, y = MM_conv, type = "l", lty = 1, col = "green", lwd =1.5)
lines(x = trials_vec, y = CV_conv, type = "l", lty = 1, col = "red", lwd = 2)

legend(x = 600, y = 0.62, legend = c("Closed Form", "Crude Monte Carlo", "Antithetic Variates"), lty = c(2,1,1), lwd =c(1.5,1.5,1.5),col = c("black","black","blue"), cex = 0.8, bty="n")

legend(x = 1200, y = 0.62, legend = c("Moment Matching","Control Variates"), lty = c(1,1), lwd =c(1.5,1.5),col = c("green","red"), cex = 0.8, bty="n")

```